---
output:
  html_notebook:
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: true
    number_sections: true  
    df_print: paged
params:
  data_files:
  kit: Biocrates AbsoluteIDQ p400 HR Kit
  measurement_type: LC
  generic_data_types:
  generic_index_first_compound:
  title: Biocrates QC Report
  author: Unknown
  profiling_variables:
  study_variables:
  replicate_variables:
  pool_indicator: Sample.Identification
  preproc_keep_status: "Valid"
  preproc_q500_urine_limits: FALSE
  filter_compound_qc_ref_max_mv_ratio: 0.3
  filter_compound_qc_ref_max_rsd: 15
  filter_compound_qc_pool_max_mv_ratio: 0.3
  filter_compound_qc_pool_max_rsd: 15
  filter_compound_bs_max_mv_ratio: 0.3
  filter_compound_bs_min_rsd: 15
  filter_sample_max_mv_ratio: 0.2
  data_tables: all
  data_export_long:
  data_export_wide:
  data_export_prefix:
  metadata_import:
  metadata_import_overlap:
  metadata_name_mods_org:
  metadata_name_mods_add:
  metadata_value_mods:
  lowcon_conditions:
  lowcon_sd_outlier_removal:
  lowcon_minimum_intensity:

  # [WIP] Optional additional normalization for biological samples (CURRENTLY DISABLED)
  # Currently either PQN or None
  # (not visibly passed via create_report function, yet)
  additional_normalization: None
  # Sample types used to create the reference sample for PQN (via median)
  # (not visibly passed via create_report function, yet)
  pqn_reference_type: Sample

  # [WIP] Directory and filename for the export of the processed dataset
  # export_dir: "."
  # export_name: "data_export"

  # [WIP] A regex string to filter/exclude samples directly after import
  sample_filter: ""

  # [WIP] Extraction of meta data from a character column
  # (not visibly passed via create_report function, yet)
  metadata_extraction:

  # Whether zeros in the data should be treated as missing values
  # (not visibly passed via create_report function, yet)
  zero2na: true
  
  # Whether report rendering should try to continue on errors
  # (not visibly passed via create_report function, switch for debugging only)
  ignore_errors: true
  
  # Enable debug mode:
  # * Export intermediary data for easy import during package development
  # (not visibly passed via create_report function, switch for debugging only)
  debug_mode: false
  
title: "`r params$title`"
author: "`r params$author`"
date: "`r format(Sys.time(), '%B %d, %Y')`"
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  error = params$ignore_errors,
  fig.width = 7,
  fig.height = 5,
  fig.align = "center"
  # out.width = "100%"
  # dev=c('png', 'pdf')
)
knitr::opts_knit$set(
  eval.after = 'fig.cap'
)
```

```{r child="nbc_setup.Rmd"}
```

This report was created with [MeTaQuaC](https://github.com/bihealth/MeTAQuaC)
v`r packageVersion("metaquac")`.

The data as imported, restructured and preprocessed in this report is available for export in the
section [Preprocessing](#preprocessing) (via the full data tables provided for each preprocessing
step).

```{r eval=IS_BIOCRATES, results='asis'}
cat("
Before reading and interpreting this QC report, please make sure to be familiar
with the Biocrates kit used, i.e. familiarize your self with the compounds,
sample types, status values, terminology, analytical specification, etc.
Please refer to Biocrates' manuals and documents provided with the kit used.\n
")
```

# Data Preparation
```{r child="nbc_import_biocrates.Rmd", eval=IS_BIOCRATES}
```

```{r child="nbc_import_generic.Rmd", eval=!IS_BIOCRATES}
```

```{r child="nbc_metadata.Rmd", eval=CONTINUE}
```

```{r child="nbc_data_restructuring.Rmd", eval=CONTINUE}
```

## Table
```{r, eval=CONTINUE}
easy_datatable(
  data = biocrates,
  caption = paste0("Full data before preprocessing")
)
```

# Overview
```{r child="nbc_import_info.Rmd", eval=CONTINUE}
```

```{r child="nbc_overview_profiling.Rmd", eval=CONTINUE}
```


# Preprocessing {.tabset #preprocessing}
```{r child="nbc_preprocessing.Rmd", eval=CONTINUE}
```


# Quality Control
```{r child="nbc_qc.Rmd", eval=CONTINUE}
```

```{r child="nbc_qc_compound_variability.Rmd", eval=CONTINUE}
```

```{r child="nbc_qc_rsd.Rmd", eval=CONTINUE}
```

```{r child="nbc_calibration_scatter.Rmd", eval=CONTINUE&&IS_BIOCRATES}
```


# Multivariate Visualization
```{r child="nbc_multivariate_visualization.Rmd", eval=CONTINUE}
```


<!-- Low Concentration Analysis (optional) -->
```{r child="nbc_misc_low_concentration.Rmd", eval=CONTINUE&&LOWCON}
```


# Miscellaneous
```{r child="nbc_misc_add_figures.Rmd", eval=CONTINUE}
```

<!-- Currently disabled, as data is available for export in the preprocessing section. -->
<!-- Might re-enable later with a parameter to allow for automated data extraction in pipelines. -->
```{r child="nbc_misc_export.Rmd", eval=FALSE}
```

```{r child="nbc_misc_environment.Rmd", eval=TRUE}
```
