## Calibration {.tabset}
Calibration scatter plots are reconstructed to enable an impression of calibration performance and
range placement (in particular for 7-point calibrated compounds). To properly reflect the internal
standard calibration method, areas and intensities are normalized by the ISTD areas and intensities,
resp.

The following visualization is based on the status-preprocessed dataset 2b.

```{r}
# Select dataset
biocrates <- datasets$filter_compounds_by_qc_rsd_kept

# Dynamic figure height
num_columns <- 7
num_compounds <- length(unique(biocrates$Compound))
compound_figure_height <- ceiling(num_compounds / num_columns) * 1 + 1
```

```{r, results="asis", fig.height=compound_figure_height}
if (meas_type == "LC"){
  cat("\n### Concentration vs peak area\n")
  plot_compound_scatter(data = biocrates,
                        x = PKG_ENV$CONCENTRATION,
                        y = "Analyte Peak Area [area]",
                        y_istd = "Internal Std. Peak Area [area]",
                        sample_types = c(SAMPLE_TYPE_BIOLOGICAL,
                                         SAMPLE_TYPE_REFERENCE_QC,
                                         SAMPLE_TYPE_POOLED_QC,
                                         paste0("Standard L", 1:7)),
                        aspect = FALSE,
                        shape = BATCH,
                        ncol = num_columns)
}
```

### Concentration vs intensity
```{r fig.height=compound_figure_height}
plot_compound_scatter(data = biocrates,
                      x = PKG_ENV$CONCENTRATION,
                      y = "Analyte Intensity [cps]",
                      y_istd = "Internal Std. Intensity [cps]",
                      sample_types = c(SAMPLE_TYPE_BIOLOGICAL,
                                       SAMPLE_TYPE_REFERENCE_QC,
                                       SAMPLE_TYPE_POOLED_QC,
                                       paste0("Standard L", 1:7)),
                      aspect = FALSE,
                      shape = BATCH,
                      ncol = num_columns)
```


```{r}
# Remove "biocrates" dataset to ensure following sections select the dataset they need
rm(biocrates)
```
