<!--
Statistical evaluation and classification of compound variability provided by Yoann Gloaguen
-->

```{r}
# Select dataset for QC
biocrates <- datasets$discarded
```

## Compound variability {.tabset}
### Overview
The plot below shows the standard deviation of each individual metabolite against their average
concentration, separately for QC and biological samples (i.e. technical vs biological variability).

A linear regression analysis is used to attempt to separate the technical and biological
variability. The outputs of the models are used to score and assess this separation, and can be used
as indicator for data quality.

```{r SD_vs_mean_model, fig.align='center'}
compound_var_stats <- compound_variability_scoring(biocrates)
```

Please see the details and the plot below for more information.

```{r SD_vs_mean_plot, fig.align='center'}
sd_mean_comp <- plot_compound_sd_vs_mean(
  data = biocrates,
  target = PKG_ENV$CONCENTRATION,
  sample_types = c(SAMPLE_TYPE_BIOLOGICAL, SAMPLE_TYPE_REFERENCE_QC),
  return_with_data = TRUE)

sd_mean_comp$scatterplot
```


### Details

<!-- A one sided Wilcoxon test is used to check that QC SDs are lower than biological sample SDs. -->

An integrative linear model is used to assess differences in technical and biological variability:
`log(SD) ~ Group + log(Mean)/Group` with Group being a factor indicating QC or biological samples.

Model results are combined to identify four important main cases of differences in technical and
biological variability:

1) `k >= 7`: A clear biological variability on top of the technical one.
2) `k = 2:6`: Inconclusive differences in biological and technical variability.
3) `k = 0:1`: A strong technical variability which hides the biological one.
4) `k < 0`: Something clearly wrong (when variability in QC decreases with increasing amount of
   metabolite concentration).

with final score k being computed to allow the classification of the four main cases as follows:

    k = 0
    k = k + (r_squared > 0.8)
    k = k + 2*(delta_intercept > 0 & delta_intercept_pvalue < 0.05)
    k = k + 4*(delta_slope > 0)
    k = k - 8*(qc_slope < 0)

The results are as follows:

* Linear regression model results:
    * R squared = `r compound_var_stats$r_squared`
    * QC slope = `r compound_var_stats$qc_slope`
    * Delta slope (Sample - QC) = `r compound_var_stats$delta_slope`
    * Delta slope p-value = `r compound_var_stats$delta_slope_pvalue`
    * Delta intercept (Sample - QC) = `r compound_var_stats$delta_intercept`
    * Delta intercept p-value = `r compound_var_stats$delta_intercept_pvalue`
* Variability Score:
    * k = `r compound_var_stats$k`

<!-- * Wilcoxon test results: -->
<!--     * Wilcoxon test p-value = `r compound_var_stats$wilcox_pvalue` -->


<!-- ### Per study variable {.tabset} -->
```{r, results="asis"}
# Compare SDs vs Means over different study variable groups
sd_plots_fun <- function(var, data, info, parent){
  # Plot compound SDs vs Means of PKG_ENV$CONCENTRATION
  plot_sd <- plot_compound_sd_vs_mean(data = data, shape = BATCH,
                                   target = PKG_ENV$CONCENTRATION,
                                   sample_types = c(SAMPLE_TYPE_BIOLOGICAL, SAMPLE_TYPE_REFERENCE_QC, SAMPLE_TYPE_POOLED_QC),
                                   study_class = var)

  plot_rsd <- plot_compound_rsd_vs_mean(data = data, shape = BATCH,
                                    target = PKG_ENV$CONCENTRATION,
                                    sample_types = c(SAMPLE_TYPE_BIOLOGICAL, SAMPLE_TYPE_REFERENCE_QC, SAMPLE_TYPE_POOLED_QC),
                                    study_class = var)

  return(list(
    Header = paste0("\n#### ", info, var, "\n"),
    Plot_sd = plot_sd,
    Plot_rsd = plot_rsd))
}

if (length(STUDY_VARIABLES) > 0) {
  cat("\n### Per study variable {.tabset}\n")
  cat("\nFor visual examination, SD as well as %RSDs are plotted against the mean per compound not
only separated by sample type but further separated by groups in study variables of the biological
samples.\n")
  var_plots <- recursive_execution(vars = STUDY_VARIABLES,
                                   end_fun = sd_plots_fun,
                                   data = biocrates,
                                   keep_sample_types = c(SAMPLE_TYPE_REFERENCE_QC, SAMPLE_TYPE_POOLED_QC))
  for (var_plot in var_plots) {
    cat(var_plot$Header)
    print(var_plot$Plot_sd)
    print(var_plot$Plot_rsd)
  }
}
```


```{r}
# Remove "biocrates" dataset to ensure following sections select the dataset they need
rm(biocrates)
```
