<!-- Author: "Yoann Gloaguen" -->
<!-- Author: "Mathias Kuhring" -->


```{r organise_data, echo=FALSE}
# Keep only columns (variable) of interest from the biocrates dataframe
selected_columns <- names(df) %in% c(
  "Compound", "Sample.Name", "Sample.Identification", "Sample.Type",
  "Analyte Intensity [cps]", "Internal Std. Intensity [cps]",
  LOWCON_CONDITIONS
)

df.filtered <- df[selected_columns]

# Renaming column for consistent naming accross datasets 
df.filtered <- df.filtered %>%
  rename(
    Analyte.Intensity = `Analyte Intensity [cps]`,
    Internal.Std.Intensity = `Internal Std. Intensity [cps]`,
  )

# Extract PBS (zero samples) first from the dataframe
df.zero_samples <- df.filtered[ df.filtered$Sample.Type == "Zero Sample", ]
# Extract QC 2 samples
df.qc_samples <- df.filtered[ df.filtered$Sample.Type == "QC Level 2", ]
# Filter the dataframe to keep only biological samples
# As we don't have any annotation file, we get our sample list directly from the dataframe
df.filtered <- df.filtered[ df.filtered$Sample.Type == "Sample", ]

# Problem in annotation, need to remove some files
# df.filtered <- df.filtered[ which(!df.filtered$Sample.Identification %in% c('44','45','47','49','50')),]
```

## Metabolites RSD in Biocrates QC samples
No filtering is done at this stage and this table is given as infromation only

```{r qc_sample_check, echo=FALSE}
# Extra step to check whether some metabolites should be removed because they have a large RSD in the QC samples

df.qc_samples.summary <- df.qc_samples %>% 
  mutate(Analyte.Peak.Intensity.Normalised = Analyte.Intensity/Internal.Std.Intensity) %>%
  group_by(Compound) %>%
  mutate(relative_standard_deviation = rsd(na.omit(Analyte.Peak.Intensity.Normalised)))

df.qc_samples.summary %>%
  distinct(Compound, .keep_all = TRUE)  %>%
  select(Compound,relative_standard_deviation) %>%
  easy_datatable()

```

```{r zero_sample_threshold, echo=FALSE}
df.zero_samples <- df.zero_samples %>%
  group_by(Compound) %>%
  mutate(mean_zero_peak_intensity = mean(Analyte.Intensity, na.rm = T), median_zero_peak_intensity = median(Analyte.Intensity, na.rm = T), highest_zero_peak_intensity = max(Analyte.Intensity, na.rm = TRUE))
# mutate(mean_peak_area = mean(Analyte.Peak.Area..area., na.rm = T), median_peak_area = median(Analyte.Peak.Area..area., na.rm = T), highest_peak_area = max(Analyte.Peak.Area..area., na.rm = TRUE), mean_concentration = mean(Concentration..??M., na.rm = T), median_concentration = median(Concentration..??M., na.rm = T), highest_concentration = max(Concentration..??M., na.rm = TRUE))

# Replace "NaN" values by NA
df.zero_samples[sapply(df.zero_samples, is.nan)] <- NA
# Replace "-Inf" values created by max() by NA
df.zero_samples[df.zero_samples=="-Inf"] <- NA

df.zero_sample_summary <- df.zero_samples %>%
  group_by(Compound) %>%
  select(Compound, mean_zero_peak_intensity,median_zero_peak_intensity) %>%
  distinct(Compound, .keep_all = T)

df.filtered <- left_join(df.filtered,df.zero_sample_summary[,c("Compound","mean_zero_peak_intensity", "median_zero_peak_intensity")], by="Compound")
```

## Number of samples above zero samples 3*median peak intensity threshold

```{r sample_threshold_calculation, echo=FALSE}

df.filtered <- df.filtered %>%
  # mutate(above_3concentration = ifelse(Concentration....M. >= mean_concentration*3, TRUE, FALSE), above_3peak_area = ifelse(Analyte.Peak.Area..area. >= mean_peak_area*3, TRUE, FALSE))
  mutate(above_3mean_peak_intensity = ifelse(Analyte.Intensity >= mean_zero_peak_intensity*1, TRUE, FALSE), above_3median_peak_intensity = ifelse(Analyte.Intensity >= median_zero_peak_intensity*1, TRUE, FALSE))

```

```{r sample_summary_table, echo=FALSE}
df.filtered %>%
  group_by(Compound) %>%
  summarise(Number_sample_above = sum(above_3median_peak_intensity == TRUE, na.rm = T), Number_sample_below = sum(above_3median_peak_intensity == FALSE, na.rm = T), Number_NA = sum(is.na(above_3median_peak_intensity))) %>% 
  easy_datatable()

```

## Number of metabolites per condition kept for analysis (after intensity and zero samples filters)
```{r condition_summary_table, echo=FALSE}
# First we set to NA all values below the threshold to discard them
df.filtered$Analyte.Intensity[ which(df.filtered$above_3median_peak_intensity == FALSE)] = NA

df.filtered <- df.filtered %>%
  mutate(Above.LOD = ifelse(Analyte.Intensity > 1000, TRUE, FALSE))

df.filtered$Analyte.Intensity[ which(df.filtered$Above.LOD == FALSE)] = NA

df.filtered <- df.filtered %>%
  mutate(Analyte.Intensity.Normalised = Analyte.Intensity/Internal.Std.Intensity)

condition_summary <- df.filtered %>%
  group_by_at(vars(Compound, one_of(LOWCON_CONDITIONS))) %>%
  summarise(Number_values = sum(!is.na(Analyte.Intensity)), Number_NA = sum(is.na(Analyte.Intensity))) %>%
  mutate(Keep = ifelse(Number_values >=1, TRUE,FALSE))

condition_summary %>%
  group_by_at(vars(one_of(LOWCON_CONDITIONS))) %>%
  summarise(Number_of_metabolite = sum(Keep == TRUE)) %>%
  easy_datatable()

```

## Total peak intensity per sample
```{r total_peak_area_per_sample_HOS, echo=FALSE, fig.height=7}
df.tot_area_HOS <- df.filtered %>%
  group_by(Sample.Name) %>%
  # filter(Cell.line == "HOS") %>%
  mutate(
    total_intensity = sum(Analyte.Intensity, na.rm = T),
    total_intensity_normalised = sum(Analyte.Intensity.Normalised, na.rm = T)
  ) %>%
  distinct(Sample.Name, .keep_all = TRUE) %>%
  ungroup()

df.tot_area_HOS$condition <- df.tot_area_HOS %>%
  select(one_of(LOWCON_CONDITIONS)) %>%
  purrr::imap(.f = ~ paste(.y, .x, sep=" ")) %>%
  as_tibble %>%
  tidyr::unite(condition, LOWCON_CONDITIONS, sep=" | ") %>%
  pull(condition)

df.tot_area_HOS$condition <- as.factor(df.tot_area_HOS$condition)

ggplot(data=df.tot_area_HOS, aes(x=reorder(Sample.Identification, as.numeric(condition)), y=total_intensity, fill=condition)) +
  geom_bar(stat="identity")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "Sample name", y = " Total peak intensity") +
  theme(legend.position = "bottom")

```

## Total peak intensity per sample after normalisation to internal standards
```{r normalised_total_peak_area_per_sample_HOS, echo=FALSE, fig.height=7}
ggplot(data=df.tot_area_HOS, aes(x=reorder(Sample.Identification, as.numeric(condition)), y=total_intensity_normalised, fill=condition)) +
  geom_bar(stat="identity")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "Sample name", y = " Total peak intensity") +
  theme(legend.position = "bottom")

# knitr::knit_exit()
```

## Outlier detection on normalised total peak intensity using a 1.5 standard deviation cutoff

Remove SD outliers (i.e. set values to NA): `r LOWCON_SD_OUTLIER_REM` (according
to parameter lowcon_sd_outlier_removal)
```{r standard_deviation_outlier_detection_HOS, echo=FALSE, fig.height=7}
# Calculate 2 standard deviation per group
df.tot_area_HOS <- df.tot_area_HOS %>%
  group_by(condition) %>%
  mutate(tot_intensity_2_sd_low = mean(total_intensity_normalised)-1.5*sd(total_intensity_normalised),tot_intensity_2_sd_high = mean(total_intensity_normalised)+1.5*sd(total_intensity_normalised)) 

df.tot_area_HOS <- df.tot_area_HOS %>%
  mutate(inside_range = ifelse(total_intensity_normalised>tot_intensity_2_sd_low & total_intensity_normalised<tot_intensity_2_sd_high, TRUE, FALSE))

ggplot(data=df.tot_area_HOS, aes(x=reorder(Sample.Identification, as.numeric(condition)), y=total_intensity_normalised, fill=condition)) +
  geom_bar(stat="identity", aes(colour=inside_range))+
  # scale_fill_manual(inside_range = c("FALSE" = "red", "TRUE" = "white"), legend = FALSE) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "Sample name", y = "Total peak intensity") +
  theme(legend.position = "bottom")

```

```{r removing_outlier_HOS, echo=FALSE}
# Remove outlier samples from the data
if (LOWCON_SD_OUTLIER_REM){
  df.tot_area_HOS$total_intensity_normalised[ which(df.tot_area_HOS$inside_range == FALSE)] = NA
}
```

## Mean peak intensity per condition (with standard deviation)

```{r mean_peak_area_per_condition_HOS, echo=FALSE, fig.height=7}

df.summary_HOS <- df.tot_area_HOS %>%
  group_by(condition) %>%
  mutate(mean_peak_intensity = mean(total_intensity_normalised, na.rm = T), standard_deviation = sd(total_intensity_normalised, na.rm = T)) %>%
  summarise(mean_peak_intensity = unique(mean_peak_intensity), standard_deviation = unique(standard_deviation))

ggplot(df.summary_HOS, aes(x=condition, y=mean_peak_intensity, fill=condition)) +
  geom_bar(position=position_dodge(), stat="identity") +
  geom_errorbar(aes(ymin=mean_peak_intensity-standard_deviation, ymax=mean_peak_intensity+standard_deviation),
                width=.2,                    # Width of the error bars
                position=position_dodge(.9)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "Condition", y = "Mean peak intensity") +
  theme(legend.position = "bottom")

```

## Summary RSD of total intensity area

```{r summary_rsd_peak_area_HOS, echo=FALSE}
summary_rsd_HOS <- df.tot_area_HOS %>%
  group_by(condition) %>%
  mutate(mean_peak_intensity = mean(total_intensity_normalised, na.rm = T), rsd = rsd(na.omit(total_intensity_normalised))) %>%
  summarise(mean_peak_intensity = unique(mean_peak_intensity), relative_standard_deviation = unique(rsd))

colnames(summary_rsd_HOS) <- c("Condition", "Mean peak intensity", "Relative Standard deviation (%)")
summary_rsd_HOS %>%
  easy_datatable()

```

## Metabolite RSD per condition

A minimum number of 3 values is required for the RSD to be reported in the table. Any condition having a metabolite in two samples or less is set to NA.  

```{r metabolite_rsd_peak_area_HOS, echo=FALSE}
# Bring condition names and inside range value to the main dataframe
df.filtered.condition.HOS <- left_join(df.filtered,df.tot_area_HOS[,c("Sample.Name","condition","inside_range")], by="Sample.Name")

# Remove outlier samples found to be outside 1.5 standard deviation
if (LOWCON_SD_OUTLIER_REM){
  df.filtered.condition.HOS$Analyte.Intensity.Normalised[ which(df.filtered.condition.HOS$inside_range == FALSE)] = NA
}

df.filtered.condition.HOS <- df.filtered.condition.HOS %>%
  # filter(Cell.line == "HOS") %>%
  group_by(condition, Compound) %>%
  mutate(relative_standard_deviation = rsd(na.omit(Analyte.Intensity.Normalised)), metabolite_mean_peak_area = mean(Analyte.Intensity.Normalised, na.rm = T), number_values=sum(!is.na(Analyte.Intensity.Normalised), na.rm = T)) %>%
  distinct(Compound, condition, .keep_all = TRUE) %>%
  mutate(relative_standard_deviation = ifelse(number_values > 2, relative_standard_deviation, NA),metabolite_mean_peak_area = ifelse(number_values > 2, metabolite_mean_peak_area, NA))

df.metabolite_rsd_HOS <- df.filtered.condition.HOS %>%
  select(Compound, condition, relative_standard_deviation) %>%
  spread(Compound, relative_standard_deviation)

df.metabolite_rsd_HOS %>%
  easy_datatable()

```

## Scatter plot of metabolites per condition (with standard error) {.tabset}

```{r HOS_scatter_plot, results="asis", echo=FALSE, fig.height=100, out.width = "100%"}
# Bring condition names and inside range value to the main dataframe
df.filtered.condition.HOS.jitter.all <- left_join(df.filtered,df.tot_area_HOS[,c("Sample.Name","condition","inside_range")], by="Sample.Name")

if (is.null(LOWCON_SCATTER_SUB_GROUPS)){
  sub_groups <- "All"
} else {
  sub_groups <- LOWCON_SCATTER_SUB_GROUPS
}

for (i in seq_along(sub_groups)) {
  if (sub_groups[i] == "All") {
    df.filtered.condition.HOS.jitter <- df.filtered.condition.HOS.jitter.all
  } else {
    df.filtered.condition.HOS.jitter <- df.filtered.condition.HOS.jitter.all %>%
      filter(!!sym(names(sub_groups[i])) == sub_groups[i])
    cat(paste("\n###", names(sub_groups[i]), sub_groups[i], "\n"))
  }
  
  # Remove outlier samples found to be outside 1.5 standard deviation
  if (LOWCON_SD_OUTLIER_REM){
    df.filtered.condition.HOS.jitter$Analyte.Intensity.Normalised[ which(df.filtered.condition.HOS.jitter$inside_range == FALSE)] = NA
  }
  
  df.filtered.condition.HOS.jitter <- df.filtered.condition.HOS.jitter %>%
    # filter(Cell.line == "HOS") %>%
    group_by(condition, Compound) %>%
    mutate(relative_standard_deviation = rsd(na.omit(Analyte.Intensity.Normalised)),standard_deviation = sd(Analyte.Intensity.Normalised, na.rm = T), metabolite_mean_peak_area = mean(Analyte.Intensity.Normalised, na.rm = T), standard_error= sd(Analyte.Intensity.Normalised, na.rm = T)/sqrt(length(na.omit(Analyte.Intensity.Normalised))), number_values=sum(!is.na(Analyte.Intensity.Normalised), na.rm = T)) %>%
    mutate(Analyte.Intensity.Normalised = ifelse(number_values > 2, Analyte.Intensity.Normalised, NA), standard_error = ifelse(number_values > 2, standard_error, NA), relative_standard_deviation = ifelse(number_values > 2, relative_standard_deviation, NA))
  
  # Get the total number of values (in samples) we have per compound
  df.filtered.condition.HOS.jitter <- df.filtered.condition.HOS.jitter %>%
    group_by(Compound) %>%
    mutate(total_number_value_available = sum(!is.na(Analyte.Intensity.Normalised), na.rm = T))
  
  # Remove all compounds with no value at all
  df.filtered.condition.HOS.jitter <- df.filtered.condition.HOS.jitter %>%
    filter(total_number_value_available > 0)
  
  # Factorize variables
  df.filtered.condition.HOS.jitter[[LOWCON_SCATTER_X]] <-
    factor(df.filtered.condition.HOS.jitter[[LOWCON_SCATTER_X]])
  if (!is.null(LOWCON_SCATTER_COLOR)){
    df.filtered.condition.HOS.jitter[[LOWCON_SCATTER_COLOR]] <-
      factor(df.filtered.condition.HOS.jitter[[LOWCON_SCATTER_COLOR]])
  } else {
    LOWCON_SCATTER_COLOR <- LOWCON_SCATTER_X
  }
  
  df.filtered.condition.HOS.jitter.error_bar <- df.filtered.condition.HOS.jitter %>%
    distinct(Compound, condition, .keep_all = TRUE)
  
  
  g <- ggplot(
    mapping = aes_string(
      x = paste0("`", LOWCON_SCATTER_X, "`"),
      y = "Analyte.Intensity.Normalised",
      color = paste0("`", LOWCON_SCATTER_COLOR, "`"),
      fill = paste0("`", LOWCON_SCATTER_COLOR, "`"))
  ) +
    geom_jitter(
      data = df.filtered.condition.HOS.jitter,
      position = position_jitterdodge(jitter.width = 0.5, dodge.width = 0.9),
      cex = 1
    ) +
    geom_errorbar(
      data = df.filtered.condition.HOS.jitter.error_bar,
      mapping = aes(
        ymin = metabolite_mean_peak_area-standard_error,
        ymax = metabolite_mean_peak_area+standard_error
      ),
      position = position_dodge(width = 0.8),
      width = 0.5, size = 0.25, color = "black"
    ) +
    facet_wrap(~ Compound, ncol = 4, scale = "free") +
    ggtitle("Peak intensity distribution per metabolite/condtion") +
    ylab("Peak intensity") +
    xlab(LOWCON_SCATTER_X) +
    theme(
      panel.grid.minor=element_blank(),
      panel.grid.major=element_blank(),
      panel.background=element_blank(),
      axis.line=element_line(),
      legend.position = "top"
    )
  
  print(g)
  
  
  if (!is.null(LOWCON_EXPORT_PATH)){
    export_path <- LOWCON_EXPORT_PATH
    dir.create(file.path(export_path), recursive = TRUE, showWarnings = TRUE)
    
    df.HOS.output.long <- df.filtered.condition.HOS.jitter %>%
      select(Compound,Sample.Identification,condition,Analyte.Intensity.Normalised,metabolite_mean_peak_area,relative_standard_deviation)
    
    # Export long format containing all data together
    write.csv(df.HOS.output.long, file.path(export_path, paste0("lowcon_", sub_groups[i], "_FIA_normalised_data_long.csv")),  row.names = FALSE)
    
    # Create a wide format with normalised intensity values
    df.HOS.output.intensity.wide <- df.HOS.output.long %>%
      mutate(sample_name = paste(Sample.Identification, condition, sep="_")) %>%
      select(Compound, sample_name, Analyte.Intensity.Normalised) %>%
      spread(sample_name, Analyte.Intensity.Normalised)
    
    write.csv(df.HOS.output.intensity.wide, file.path(export_path, paste0("lowcon_", sub_groups[i], "_FIA_normalised_intensities_wide.csv")), row.names = FALSE)
    
    # Create a wide format with mean and RSD values for each condition
    df.HOS.output.mean.wide <- df.HOS.output.long %>%
      # mutate(sample_name = paste(Sample.Identification, Volume, Mass, BR, sep="_")) %>%
      select(Compound, condition, metabolite_mean_peak_area) %>%
      distinct(Compound, condition, .keep_all = TRUE) %>%
      spread(condition, metabolite_mean_peak_area)
    
    write.csv(df.HOS.output.mean.wide, file.path(export_path, paste0("lowcon_", sub_groups[i], "_FIA_mean_intensity_wide.csv")), row.names = FALSE)
    # 
    df.HOS.output.rsd.wide <- df.HOS.output.long %>%
      # mutate(sample_name = paste(Sample.Identification, Volume, Mass, BR, sep="_")) %>%
      select(Compound, condition, relative_standard_deviation) %>%
      distinct(Compound, condition, .keep_all = TRUE) %>%
      spread(condition, relative_standard_deviation)
    
    write.csv(df.HOS.output.rsd.wide, file.path(export_path, paste0("lowcon_", sub_groups[i], "_FIA_rsd_wide.csv")), row.names = FALSE)
  }
}

```
