## Compound %RSDs

### Replicate %RSDs
Relative standard deviations per compound or compound class are calculated and visualized
for different sample groupings, either by sample type or for biological replicates (i.e.
samples from the same group within variables indicated with parameter
`replicate_variables`) as well as for all batches or each batch separately.

This analysis is based on the status-preprocessed dataset 1.

```{r}
# Select dataset
biocrates <- datasets$discarded
```

```{r}
# Figure size calculations
calc_fig_sizes_rep_rsd <- function(bcdata, rep_variables, sample_types, unit, legends = 0,
                                   num_combinations = NULL, unit_factor = 1){
  num_variables <- length(rep_variables)
  num_units <- bcdata %>% pull(UQ(sym(unit))) %>% unique() %>% length()
  if (is.null(num_combinations)){
    num_combinations <- bcdata %>%
      filter(Sample.Type %in% sample_types) %>%
      select(one_of(rep_variables)) %>% distinct() %>% nrow()
  }

  facet_parts <- max(unit_factor * num_units + 3.5 * num_variables,
                     3.5 * num_variables * 2.5)
  facet_parts_max <- 90

  num_columns <- min(max(floor(facet_parts_max / facet_parts), 1), num_combinations)
  num_rows <- ceiling(num_combinations/num_columns)

  fig_width_rep_rsd <- 7
  if (facet_parts_max / facet_parts < 1){
    fig_width_rep_rsd <- (1/7) + (6/7) * 7 / facet_parts_max * facet_parts
  }
  fig_height_rep_rsd <- max((num_rows + 1) * 1.2, 3.5) + legends

  return(list(h = fig_height_rep_rsd, w = fig_width_rep_rsd, c = num_columns))
}
```

#### Compounds {.tabset}
##### All batches {.tabset}
###### Sample types
```{r}
rep_variables <- c("Sample.Type")
sample_types <- c(SAMPLE_TYPE_POOLED_QC, SAMPLE_TYPE_REFERENCE_QC, SAMPLE_TYPE_BIOLOGICAL)
# Calculate figure size
fig_sizes_rep_rsd <- calc_fig_sizes_rep_rsd(
  biocrates, rep_variables, sample_types, "Compound")
```

```{r, fig.height=fig_sizes_rep_rsd$h, fig.width=fig_sizes_rep_rsd$w}
plot_bio_replicate_rsd(bcdata = biocrates, rep_variables = rep_variables,
                       sample_types = sample_types,
                       facet_cols = fig_sizes_rep_rsd$c)
```

```{r}
#### Biological replicates
if (length(REPLICATE_VARIABLES) > 0){
  rep_variables <- REPLICATE_VARIABLES
  sample_types <- c(SAMPLE_TYPE_BIOLOGICAL)
  # Calculate figure size
  fig_sizes_rep_rsd <- calc_fig_sizes_rep_rsd(
    biocrates, rep_variables, sample_types, "Compound")
}
```

```{r, results="asis", fig.height=fig_sizes_rep_rsd$h, fig.width=fig_sizes_rep_rsd$w}
if (length(REPLICATE_VARIABLES) > 0){
  cat(paste0("\n###### Study groups \n"))
  print(plot_bio_replicate_rsd(bcdata = biocrates, rep_variables = rep_variables,
                               sample_types = sample_types,
                               facet_cols = fig_sizes_rep_rsd$c))

  if (all(c("BR", "TR") %in% names(biocrates))){
    cat(paste0("\n###### BRs (mean TRs)\n"))
    print(plot_bio_replicate_rsd(bcdata = biocrates, rep_variables = rep_variables,
                                 sample_types = sample_types, tec_reps_use = "mean",
                                 facet_cols = fig_sizes_rep_rsd$c))

    cat(paste0("\n###### BRs (median TRs)\n"))
    print(plot_bio_replicate_rsd(bcdata = biocrates, rep_variables = rep_variables,
                                 sample_types = sample_types, tec_reps_use = "median",
                                 facet_cols = fig_sizes_rep_rsd$c))
  }
}
```

```{r}
#### Technical replicates
if (length(REPLICATE_VARIABLES) > 0){
  if (all(c("BR", "TR") %in% names(biocrates))){
    rep_variables <- REPLICATE_VARIABLES
    sample_types <- c(SAMPLE_TYPE_BIOLOGICAL)
    num_combinations <- biocrates %>%
      filter(Sample.Type %in% sample_types) %>%
      select(one_of(rep_variables), BR, TR) %>% distinct() %>%
      group_by_at(vars(one_of(rep_variables), BR)) %>%
      filter(all(n() > 1)) %>% ungroup() %>%
      select(one_of(rep_variables)) %>% distinct() %>%  nrow()
    max_tr_number <-  biocrates %>%
      filter(Sample.Type %in% sample_types) %>%
      select(one_of(rep_variables), BR, TR) %>% distinct() %>%
      group_by_at(vars(one_of(rep_variables), BR)) %>%
      filter(all(n() > 1)) %>% summarize(Number = n())
    max_tr_number <- max(max_tr_number$Number)
    unit_factor <- max(1, max_tr_number/2)
    # Calculate figure size
    fig_sizes_rep_rsd <- calc_fig_sizes_rep_rsd(
      biocrates, c(rep_variables), sample_types, "Compound", legends = 1,
      num_combinations = num_combinations, unit_factor = unit_factor)
  }
}
```

```{r, results="asis", fig.height=fig_sizes_rep_rsd$h, fig.width=fig_sizes_rep_rsd$w}
if (length(REPLICATE_VARIABLES) > 0){
  if (all(c("BR", "TR") %in% names(biocrates))){
    cat(paste0("\n###### Technical replicates only\n"))
    print(plot_bio_replicate_rsd(bcdata = biocrates, rep_variables = rep_variables,
                                 sample_types = sample_types, tec_reps_use = "only",
                                 facet_cols = fig_sizes_rep_rsd$c))
  }
}
```


##### Per batch {.tabset}
###### Sample types
```{r}
rep_variables <- c("Batch", "Sample.Type")
sample_types <- c(SAMPLE_TYPE_POOLED_QC, SAMPLE_TYPE_REFERENCE_QC, SAMPLE_TYPE_BIOLOGICAL)
# Calculate figure size
fig_sizes_rep_rsd <- calc_fig_sizes_rep_rsd(
  biocrates, rep_variables, sample_types, "Compound")
```

```{r, fig.height=fig_sizes_rep_rsd$h, fig.width=fig_sizes_rep_rsd$w}
plot_bio_replicate_rsd(bcdata = biocrates, rep_variables = rep_variables,
                       sample_types = sample_types,
                       facet_cols = fig_sizes_rep_rsd$c)
```

```{r}
#### Biological replicates
if (length(REPLICATE_VARIABLES) > 0){
  rep_variables <- c("Batch", REPLICATE_VARIABLES)
  sample_types <- c(SAMPLE_TYPE_BIOLOGICAL)
  # Calculate figure size
  fig_sizes_rep_rsd <- calc_fig_sizes_rep_rsd(
    biocrates, rep_variables, sample_types, "Compound")
}
```

```{r, results="asis", fig.height=fig_sizes_rep_rsd$h, fig.width=fig_sizes_rep_rsd$w}
if (length(REPLICATE_VARIABLES) > 0){
  cat(paste0("\n###### Study groups \n"))
  print(plot_bio_replicate_rsd(bcdata = biocrates, rep_variables = rep_variables,
                               sample_types = sample_types,
                               facet_cols = fig_sizes_rep_rsd$c))

  if (all(c("BR", "TR") %in% names(biocrates))){
    cat(paste0("\n###### BRs (mean TRs)\n"))
    print(plot_bio_replicate_rsd(bcdata = biocrates, rep_variables = rep_variables,
                                 sample_types = sample_types, tec_reps_use = "mean",
                                 facet_cols = fig_sizes_rep_rsd$c))

    cat(paste0("\n###### BRs (median TRs)\n"))
    print(plot_bio_replicate_rsd(bcdata = biocrates, rep_variables = rep_variables,
                                 sample_types = sample_types, tec_reps_use = "median",
                                 facet_cols = fig_sizes_rep_rsd$c))
  }
}
```

```{r}
#### Technical replicates
if (length(REPLICATE_VARIABLES) > 0){
  if (all(c("BR", "TR") %in% names(biocrates))){
    rep_variables <- REPLICATE_VARIABLES
    sample_types <- c(SAMPLE_TYPE_BIOLOGICAL)
    num_combinations <- biocrates %>%
      filter(Sample.Type %in% sample_types) %>%
      select(one_of(rep_variables), BR, TR) %>% distinct() %>%
      group_by_at(vars(one_of(rep_variables), BR)) %>%
      filter(all(n() > 1)) %>% ungroup() %>%
      select(one_of(rep_variables)) %>% distinct() %>%  nrow()
    max_tr_number <-  biocrates %>%
      filter(Sample.Type %in% sample_types) %>%
      select(one_of(rep_variables), BR, TR) %>% distinct() %>%
      group_by_at(vars(one_of(rep_variables), BR)) %>%
      filter(all(n() > 1)) %>% summarize(Number = n())
    max_tr_number <- max(max_tr_number$Number)
    unit_factor <- max(1, max_tr_number/2)
    # Calculate figure size
    fig_sizes_rep_rsd <- calc_fig_sizes_rep_rsd(
      biocrates, c(rep_variables), sample_types, "Compound", legends = 1,
      num_combinations = num_combinations, unit_factor = unit_factor)
  }
}
```

```{r, results="asis", fig.height=fig_sizes_rep_rsd$h, fig.width=fig_sizes_rep_rsd$w}
if (length(REPLICATE_VARIABLES) > 0){
  if (all(c("BR", "TR") %in% names(biocrates))){
    cat(paste0("\n###### Technical replicates only\n"))
    print(plot_bio_replicate_rsd(bcdata = biocrates, rep_variables = rep_variables,
                                 sample_types = sample_types, tec_reps_use = "only",
                                 facet_cols = fig_sizes_rep_rsd$c))
  }
}
```


#### Classes {.tabset}
##### All batches {.tabset}
###### Sample types
```{r}
rep_variables <- c("Sample.Type")
sample_types <- c(SAMPLE_TYPE_POOLED_QC, SAMPLE_TYPE_REFERENCE_QC, SAMPLE_TYPE_BIOLOGICAL)
# Calculate figure size
fig_sizes_rep_rsd <- calc_fig_sizes_rep_rsd(
  biocrates, rep_variables, sample_types, "Class")
```

```{r, fig.height=fig_sizes_rep_rsd$h, fig.width=fig_sizes_rep_rsd$w}
plot_bio_replicate_rsd(bcdata = biocrates, rep_variables = rep_variables,
                       sample_types = sample_types,
                       facet_cols = fig_sizes_rep_rsd$c, unit = "Class")
```

```{r}
###### Biological replicates
if (length(REPLICATE_VARIABLES) > 0){
  rep_variables <- REPLICATE_VARIABLES
  sample_types <- c(SAMPLE_TYPE_BIOLOGICAL)
  # Calculate figure size
  fig_sizes_rep_rsd <- calc_fig_sizes_rep_rsd(
    biocrates, rep_variables, sample_types, "Class")
}
```

```{r, results="asis", fig.height=fig_sizes_rep_rsd$h, fig.width=fig_sizes_rep_rsd$w}
if (length(REPLICATE_VARIABLES) > 0){
  cat("\n###### Study groups\n")
  print(plot_bio_replicate_rsd(bcdata = biocrates, rep_variables = rep_variables,
                               sample_types = sample_types,
                               facet_cols = fig_sizes_rep_rsd$c, unit = "Class"))

  if (all(c("BR", "TR") %in% names(biocrates))){
    cat("\n###### BRs (mean TRs)\n")
    print(plot_bio_replicate_rsd(bcdata = biocrates, rep_variables = rep_variables,
                                 sample_types = sample_types, tec_reps_use = "mean",
                                 facet_cols = fig_sizes_rep_rsd$c, unit = "Class"))

    cat("\n###### BRs (median TRs)\n")
    print(plot_bio_replicate_rsd(bcdata = biocrates, rep_variables = rep_variables,
                                 sample_types = sample_types, tec_reps_use = "median",
                                 facet_cols = fig_sizes_rep_rsd$c, unit = "Class"))
  }
}
```

```{r}
#### Technical replicates
if (length(REPLICATE_VARIABLES) > 0){
  if (all(c("BR", "TR") %in% names(biocrates))){
    rep_variables <- REPLICATE_VARIABLES
    sample_types <- c(SAMPLE_TYPE_BIOLOGICAL)
    num_combinations <- biocrates %>%
      filter(Sample.Type %in% sample_types) %>%
      select(one_of(rep_variables), BR, TR) %>% distinct() %>%
      group_by_at(vars(one_of(rep_variables), BR)) %>%
      filter(all(n() > 1)) %>% ungroup() %>%
      select(one_of(rep_variables)) %>% distinct() %>%  nrow()
    max_tr_number <-  biocrates %>%
      filter(Sample.Type %in% sample_types) %>%
      select(one_of(rep_variables), BR, TR) %>% distinct() %>%
      group_by_at(vars(one_of(rep_variables), BR)) %>%
      filter(all(n() > 1)) %>% summarize(Number = n())
    max_tr_number <- max(max_tr_number$Number)
    unit_factor <- max(1, max_tr_number/2)
    # Calculate figure size
    fig_sizes_rep_rsd <- calc_fig_sizes_rep_rsd(
      biocrates, c(rep_variables), sample_types, "Class", legends = 1,
      num_combinations = num_combinations, unit_factor = unit_factor)
  }
}
```

```{r, results="asis", fig.height=fig_sizes_rep_rsd$h, fig.width=fig_sizes_rep_rsd$w}
if (length(REPLICATE_VARIABLES) > 0){
  if (all(c("BR", "TR") %in% names(biocrates))){
    cat("\n###### Technical replicates only\n")
    print(plot_bio_replicate_rsd(bcdata = biocrates, rep_variables = rep_variables,
                                 sample_types = sample_types, tec_reps_use = "only",
                                 facet_cols = fig_sizes_rep_rsd$c, unit = "Class"))
  }
}
```


##### Per batch {.tabset}
###### Sample types
```{r}
rep_variables <- c("Batch", "Sample.Type")
sample_types <- c(SAMPLE_TYPE_POOLED_QC, SAMPLE_TYPE_REFERENCE_QC, SAMPLE_TYPE_BIOLOGICAL)
# Calculate figure size
fig_sizes_rep_rsd <- calc_fig_sizes_rep_rsd(
  biocrates, rep_variables, sample_types, "Class")
```

```{r, fig.height=fig_sizes_rep_rsd$h, fig.width=fig_sizes_rep_rsd$w}
plot_bio_replicate_rsd(bcdata = biocrates, rep_variables = rep_variables,
                       sample_types = sample_types,
                       facet_cols = fig_sizes_rep_rsd$c, unit = "Class")
```

```{r}
###### Biological replicates
if (length(REPLICATE_VARIABLES) > 0){
  rep_variables <- c("Batch", REPLICATE_VARIABLES)
  sample_types <- c(SAMPLE_TYPE_BIOLOGICAL)
  # Calculate figure size
  fig_sizes_rep_rsd <- calc_fig_sizes_rep_rsd(
    biocrates, rep_variables, sample_types, "Class")
}
```

```{r, results="asis", fig.height=fig_sizes_rep_rsd$h, fig.width=fig_sizes_rep_rsd$w}
if (length(REPLICATE_VARIABLES) > 0){
  cat("\n###### Study groups\n")
  print(plot_bio_replicate_rsd(bcdata = biocrates, rep_variables = rep_variables,
                               sample_types = sample_types,
                               facet_cols = fig_sizes_rep_rsd$c, unit = "Class"))

  if (all(c("BR", "TR") %in% names(biocrates))){
    cat("\n###### BRs (mean TRs)\n")
    print(plot_bio_replicate_rsd(bcdata = biocrates, rep_variables = rep_variables,
                                 sample_types = sample_types, tec_reps_use = "mean",
                                 facet_cols = fig_sizes_rep_rsd$c, unit = "Class"))

    cat("\n###### BRs (median TRs)\n")
    print(plot_bio_replicate_rsd(bcdata = biocrates, rep_variables = rep_variables,
                                 sample_types = sample_types, tec_reps_use = "median",
                                 facet_cols = fig_sizes_rep_rsd$c, unit = "Class"))
  }
}
```

```{r}
#### Technical replicates
if (length(REPLICATE_VARIABLES) > 0){
  if (all(c("BR", "TR") %in% names(biocrates))){
    rep_variables <- REPLICATE_VARIABLES
    sample_types <- c(SAMPLE_TYPE_BIOLOGICAL)
    num_combinations <- biocrates %>%
      filter(Sample.Type %in% sample_types) %>%
      select(one_of(rep_variables), BR, TR) %>% distinct() %>%
      group_by_at(vars(one_of(rep_variables), BR)) %>%
      filter(all(n() > 1)) %>% ungroup() %>%
      select(one_of(rep_variables)) %>% distinct() %>%  nrow()
    max_tr_number <-  biocrates %>%
      filter(Sample.Type %in% sample_types) %>%
      select(one_of(rep_variables), BR, TR) %>% distinct() %>%
      group_by_at(vars(one_of(rep_variables), BR)) %>%
      filter(all(n() > 1)) %>% summarize(Number = n())
    max_tr_number <- max(max_tr_number$Number)
    unit_factor <- max(1, max_tr_number/2)
    # Calculate figure size
    fig_sizes_rep_rsd <- calc_fig_sizes_rep_rsd(
      biocrates, c(rep_variables), sample_types, "Class", legends = 1,
      num_combinations = num_combinations, unit_factor = unit_factor)
  }
}
```

```{r, results="asis", fig.height=fig_sizes_rep_rsd$h, fig.width=fig_sizes_rep_rsd$w}
if (length(REPLICATE_VARIABLES) > 0){
  if (all(c("BR", "TR") %in% names(biocrates))){
    cat("\n###### Technical replicates only\n")
    print(plot_bio_replicate_rsd(bcdata = biocrates, rep_variables = rep_variables,
                                 sample_types = sample_types, tec_reps_use = "only",
                                 facet_cols = fig_sizes_rep_rsd$c, unit = "Class"))
  }
}
```


```{r}
# Remove "biocrates" dataset to ensure following sections select the dataset they need
rm(biocrates)
```
