## Additional figures

### Replicate dispersion {.tabset}
The following figures illustrate reproducibility by comparing compound measures
(e.g. concentration) within and between groups such as different sample types and
conditions in study variables (as indicated with parameter `study_variables`).

This analysis is based on status-preprocessed dataset 3a.

```{r}
# Select dataset
biocrates <- datasets$filter_compounds_by_sample_all_mv_kept
```

```{r}
num_compounds <- biocrates %>% pull(Compound) %>% unique() %>% length()
num_columns <- 3
fig_height_dispersion <- ceiling(num_compounds/num_columns) * 1 + 1.5
```

#### Per sample type
```{r, fig.height=fig_height_dispersion}
# Illustrate concentrations in different sample types
plot_replicate_dispersion(bcdata = biocrates,
                          grouping = "Sample.Type",
                          color = "Sample.Type")
```

#### Per study variable {.tabset}
```{r, results="asis", fig.height=fig_height_dispersion}
# Concentration dispersion plots over different study variable groups
disp_plots_fun <- function(var, data, info, parent = NULL, name){
  if (is.null(parent)){
    plot <- plot_replicate_dispersion(bcdata = data,
                                      sample_types = SAMPLE_TYPE_BIOLOGICAL,
                                      grouping = var,
                                      color = var)
  } else {
    plot <- plot_replicate_dispersion(bcdata = data,
                                      sample_types = SAMPLE_TYPE_BIOLOGICAL,
                                      grouping = var,
                                      color = parent)
  }

  return(list(
    Header = paste0("\n##### ", info, var, "\n"),
    Plot = plot))
}

if (length(STUDY_VARIABLES) > 0){
  var_plots <- recursive_execution(vars = STUDY_VARIABLES,
                                   end_fun = disp_plots_fun,
                                   data = biocrates,
                                   separate_parent = FALSE)
  for (var_plot in var_plots){
    cat(var_plot$Header)
    print(var_plot$Plot)
  }
}
```


### Concentration Profiles {.tabset}
Simple metabolite concentration profiles overlapped per sample (point and line plot) or summarized
(as box plots) are created to enable a brief impression of overall sample behavior consistency (e.g.
to justify later normalization).

This analysis is based on the status-preprocessed dataset 1.

```{r}
# Select dataset
biocrates <- datasets$discarded
```

```{r}
rep_variables <- c("Sample.Name")
sample_types <- c(SAMPLE_TYPE_BIOLOGICAL) # just need to indicate one type
# Calculate figure sizes
fig_sizes_points <- calc_fig_sizes_rep_rsd(
  biocrates, rep_variables, sample_types, "Compound", num_combinations = 1)
fig_sizes_box <- calc_fig_sizes_rep_rsd(
  biocrates, rep_variables, sample_types, "Compound", num_combinations = 1, legends = 1)
```


```{r, fig.height=fig_sizes_points$h, fig.width=fig_sizes_points$w, eval=AVAILABLE_REFERENCE_QC, results="asis"}
cat(paste0("\n#### Reference QCs\n"))
plot_sample_profiles_points(
  bcdata = biocrates, target = PKG_ENV$CONCENTRATION, sample_types = SAMPLE_TYPE_REFERENCE_QC)
```

```{r, fig.height=fig_sizes_box$h, fig.width=fig_sizes_box$w, eval=AVAILABLE_REFERENCE_QC}
plot_sample_profiles_box(
  bcdata = biocrates, target = PKG_ENV$CONCENTRATION, sample_types = SAMPLE_TYPE_REFERENCE_QC)
```


```{r, fig.height=fig_sizes_points$h, fig.width=fig_sizes_points$w, eval=AVAILABLE_POOLED_QC, results="asis"}
cat(paste0("\n#### ", SAMPLE_TYPE_POOLED_QC, "s\n"))
plot_sample_profiles_points(
  bcdata = biocrates, target = PKG_ENV$CONCENTRATION, sample_types = SAMPLE_TYPE_POOLED_QC)
```

```{r, fig.height=fig_sizes_box$h, fig.width=fig_sizes_box$w, eval=AVAILABLE_POOLED_QC}
plot_sample_profiles_box(
  bcdata = biocrates, target = PKG_ENV$CONCENTRATION, sample_types = SAMPLE_TYPE_POOLED_QC)
```


#### Biological samples
```{r, fig.height=fig_sizes_points$h, fig.width=fig_sizes_points$w}
plot_sample_profiles_points(
  bcdata = biocrates, target = PKG_ENV$CONCENTRATION, sample_types = SAMPLE_TYPE_BIOLOGICAL)
```

```{r, fig.height=fig_sizes_box$h, fig.width=fig_sizes_box$w}
plot_sample_profiles_box(
  bcdata = biocrates, target = PKG_ENV$CONCENTRATION, sample_types = SAMPLE_TYPE_BIOLOGICAL)
```


```{r, fig.height=fig_sizes_points$h, fig.width=fig_sizes_points$w, eval=NORMALIZED, results='asis'}
cat("#### Biological samples (normalized)\n")
plot_sample_profiles_points(
  bcdata = biocrates_normalized, target = PKG_ENV$CONCENTRATION, sample_types = SAMPLE_TYPE_BIOLOGICAL)
```

```{r, fig.height=fig_sizes_box$h, fig.width=fig_sizes_box$w, eval=NORMALIZED}
plot_sample_profiles_box(
  bcdata = biocrates_normalized, target = PKG_ENV$CONCENTRATION, sample_types = SAMPLE_TYPE_BIOLOGICAL)
```


```{r}
# Remove "biocrates" dataset to ensure following sections select the dataset they need
rm(biocrates)
```
