```{r}
# Preprocessing parameters
ppparams <- list(
  statuses_to_keep = params$preproc_keep_status,
  statuses_q500_urine = PREPROC_Q500_URINE_LIMITS,
  threshold_2a = params$filter_compound_qc_ref_max_mv_ratio,
  threshold_2b = params$filter_compound_qc_ref_max_rsd,
  threshold_2c = params$filter_compound_qc_pool_max_mv_ratio,
  threshold_2d = params$filter_compound_qc_pool_max_rsd,
  threshold_3a = 1,
  threshold_3b = params$filter_compound_bs_max_mv_ratio,
  threshold_3c = params$filter_compound_bs_min_rsd,
  threshold_4  = params$filter_sample_max_mv_ratio
)
```

The following preprocessing procedures are applied on the data in this order:

1. **Status Preprocessing**
    a. Discard unreliable measurements based on status
2. **Remove unreliable __compounds__** (QC-based filters)
    a. Based on missing values ratio in Reference QCs (**>=`r ppparams$threshold_2a * 100`%**)
    b. Based on %RSD of Reference QCs (**>=`r ppparams$threshold_2b`%**)
    c. Based on missing values ratio in Pooled QCs (**>=`r ppparams$threshold_2c * 100`%**)
    d. Based on %RSD of Pooled QCs (**>=`r ppparams$threshold_2d`%**)
3. **Remove underrepresented or invariable __compounds__** (biological filters)
    a. Based on only missing values in biological samples (**=`r ppparams$threshold_3a * 100`%**)
    b. Based on missing values ratio in biological samples (**>=`r ppparams$threshold_3b * 100`%**)
    c. Based on %RSD in biological samples (**<`r ppparams$threshold_3c`%**)
4. **Remove underrepresented __biological samples__** (biological filters)
    a. Based on missing values ratio of compounds (**>=`r ppparams$threshold_4 * 100`%**)

The preprocessing statuses and filter thresholds can be modified with the following parameters of
the `create_report` function (refer to the readme or function description for details):

* `preproc_keep_status`
* `filter_compound_qc_ref_max_mv_ratio`
* `filter_compound_qc_ref_max_rsd`
* `filter_compound_qc_pool_max_mv_ratio`
* `filter_compound_qc_pool_max_rsd`
* `filter_compound_bs_max_mv_ratio`
* `filter_compound_bs_min_rsd`
* `filter_sample_max_mv_ratio`

The result of each preprocessing or filter step can be found in the tab with the corresponding
enumerator. Please note, as each tab contains a table with the complete dataset in the
corresponding filtered form, loading the tab for the first time might take a couple of seconds.

```{r}
message("**Note**: The beginning of each following section will indicate on which dataset
calculations and visualizations are based. Thereby, the preprocessed datasets will be
referenced as **1**, **2a**, **2b**, **3a**, **3b** and **4**, resp.")

# Execute preprocessing
datasets <- execute_preprocessing(data = biocrates, ppparams = ppparams)
# Remove "biocrates" dataset to ensure following sections select the preprocessed dataset they need
rm(biocrates)
```


## Step **1**

1. **Status Preprocessing**
    a. Discard unreliable measurements based on status

The status preprocessing discards measurements based on quality statuses.

By default, only "Valid" and "Semi Quant." measurements are considered as reliable and are kept
for further analysis. Hence, all other statuses including "< LLOQ" and "> ULOQ"
(out of quantification range / extrapolated calibration), "STD/QC < Limit" and
"STD/QC > Limit" (insufficient accuracy in STDs and QCs), "ISTD Out of Range"
(internal standard too high or low) and other statuses are usually considered as
unreliable and measurements are transformed to missing values (set to NA).

```{r eval=!IS_BIOCRATES, results="asis"}
cat("Note: Status names may differ in custom generic data.\n")
```

```{r eval=PREPROC_Q500_URINE_LIMITS, results="asis"}
message("Note: Statuses \"< LLOQ\" and \"> ULOQ\" have been updated if data is
within range of cal 0.25 and cal 9, resp. (i.e. set to \"Valid\", see parameter
`preproc_q500_urine_limits`).\n")
```

Based on the parameter `preproc_keep_status`, measurements with the following statuses will be
retained:
**`r paste0(ppparams$statuses_to_keep, collapse = "**, **")`**

Missing **Concentration** values before status preprocessing:
**`r sum(is.na(datasets$original[[ENV$CONCENTRATION]]))`** of **`r nrow(datasets$original)`**
(`r round(sum(is.na(datasets$original[[ENV$CONCENTRATION]])) / nrow(datasets$original) * 100)`%)

Missing **Concentration** values after status preprocessing:
**`r sum(is.na(datasets$discarded[[ENV$CONCENTRATION]]))`** of **`r nrow(datasets$discarded)`**
(`r round(sum(is.na(datasets$discarded[[ENV$CONCENTRATION]])) / nrow(datasets$discarded) * 100)`%)

```{r}
easy_datatable(
  summarize_missing_values_per_sample(datasets$discarded, plot = FALSE),
  round_colums = "% Missing Values", show_type = "statistics",
  caption = "Missing values per sample after status preprocessing")

easy_datatable(
  summarize_missing_values_per_compound(datasets$discarded, plot = FALSE),
  round_colums = "% Missing Values", show_type = "statistics",
  caption = "Missing values per compound after status preprocessing")

easy_datatable(
  datasets$discarded,
  caption = paste0("Full data after status preprocessing"))
```

```{r eval=PREPROC_Q500_URINE_LIMITS}
easy_datatable(
  datasets$original %>%
    select(all_of(COLUMN_SAMPLE_NAME), Compound, MetIDQ_Status) %>%
    rename(Status_before = MetIDQ_Status) %>%
    left_join(
      datasets$original_urine %>%
        select(all_of(COLUMN_SAMPLE_NAME), Compound, MetIDQ_Status) %>%
        rename(Status_after = MetIDQ_Status),
      by = c(COLUMN_SAMPLE_NAME, "Compound")
    ) %>%
    filter(Status_before != Status_after),
  show_type = "statistics",
  caption = "Status values changed according due Q500 urine cal limits")
```


## Step **2a**

2. **Remove unreliable __compounds__** (QC-based filters)
    a. Based on missing values ratio in Reference QCs (**>=`r ppparams$threshold_2a * 100`%**)

To alter the threshold use  parameter `filter_compound_qc_ref_max_mv_ratio`.

Number of compounds before: **`r length(unique(datasets$discarded$Compound))`**
Number of compounds left: **`r length(unique(datasets$filter_compounds_by_qc_mv_kept$Compound))`**

```{r}
# 2. Remove unreliable compounds
#   a. Based on missing values ratio in QCs (samples of type ENV$SAMPLE_TYPE_REFERENCE_QC)
#      Valid ratio of NAs is max 0.2 by default
easy_datatable(
  datasets$filter_compounds_by_qc_mv_removed, show_type = "statistics",
  caption = paste0("Removed compounds due to ", ppparams$threshold_2a*100, "% QC MVs"))
easy_datatable(
  datasets$filter_compounds_by_qc_mv_kept,
  caption = paste0("Full data kept after ", ppparams$threshold_2a*100, "% QC MVs filtering"))
```


## Step **2b**

2. **Remove unreliable __compounds__** (QC-based filters)
    b. Based on %RSD of Reference QCs (**>=`r ppparams$threshold_2b`%**)

To alter the threshold use parameter `filter_compound_qc_ref_max_rsd`.

Number of compounds before: **`r length(unique(datasets$filter_compounds_by_qc_mv_kept$Compound))`**
Number of compounds left: **`r length(unique(datasets$filter_compounds_by_qc_rsd_kept$Compound))`**

```{r}
easy_datatable(
  datasets$filter_compounds_by_qc_rsd_removed, show_type = "statistics",
  caption = paste0("Removed compounds due to ", ppparams$threshold_2b, " %RSD in QCs"))
easy_datatable(
  datasets$filter_compounds_by_qc_rsd_kept,
  caption = paste0("Full data kept after ", ppparams$threshold_2b, "%RSD QC filtering"))
```


## Step **2c**

2. **Remove unreliable __compounds__** (QC-based filters)
    c. Based on missing values ratio in Pooled QCs (**>=`r ppparams$threshold_2c * 100`%**)

To alter the threshold use  parameter `filter_compound_qc_pool_max_mv_ratio`.

Number of compounds before: **`r length(unique(datasets$filter_compounds_by_qc_rsd_kept$Compound))`**
Number of compounds left: **`r length(unique(datasets$filter_compounds_by_qc_pool_mv_kept$Compound))`**

```{r}
# 2. Remove unreliable compounds
#   a. Based on missing values ratio in QCs (samples of type ENV$SAMPLE_TYPE_REFERENCE_QC)
#      Valid ratio of NAs is max 0.2 by default
easy_datatable(
  datasets$filter_compounds_by_qc_pool_mv_removed, show_type = "statistics",
  caption = paste0("Removed compounds due to ", ppparams$threshold_2c*100, "% Pooled QC MVs"))
easy_datatable(
  datasets$filter_compounds_by_qc_pool_mv_kept,
  caption = paste0("Full data kept after ", ppparams$threshold_2c*100, "% Pooled QC MVs filtering"))
```


## Step **2d**

2. **Remove unreliable __compounds__** (QC-based filters)
    d. Based on %RSD of Pooled QCs (**>=`r ppparams$threshold_2d`%**)

To alter the threshold use parameter `filter_compound_qc_pool_max_rsd`.

Number of compounds before: **`r length(unique(datasets$filter_compounds_by_qc_pool_mv_kept$Compound))`**
Number of compounds left: **`r length(unique(datasets$filter_compounds_by_qc_pool_rsd_kept$Compound))`**

```{r}
easy_datatable(
  datasets$filter_compounds_by_qc_pool_rsd_removed, show_type = "statistics",
  caption = paste0("Removed compounds due to ", ppparams$threshold_2d, " %RSD in Pooled QCs"))
easy_datatable(
  datasets$filter_compounds_by_qc_pool_rsd_kept,
  caption = paste0("Full data kept after ", ppparams$threshold_2d, "%RSD Pooled QC filtering"))
```


## Step **3a**

3. **Remove underrepresented __compounds__** (biological filters)
    a. Based on only missing values in biological samples (**=`r ppparams$threshold_3a * 100`%**)

Number of compounds before: **`r length(unique(datasets$filter_compounds_by_qc_pool_rsd_kept$Compound))`**
Number of compounds left: **`r length(unique(datasets$filter_compounds_by_sample_all_mv_kept$Compound))`**

```{r}
easy_datatable(
  datasets$filter_compounds_by_sample_all_mv_removed, show_type = "statistics",
  caption = paste0("Removed compounds due to ", ppparams$threshold_3a*100, "% Sample MVs"))
easy_datatable(
  datasets$filter_compounds_by_sample_all_mv_kept,
  caption = paste0("Full data kept after ", ppparams$threshold_3a*100, "% Sample MVs filtering"))
```


## Step **3b**

3. **Remove underrepresented __compounds__** (biological filters)
    b. Based on missing values ratio in biological samples (**>=`r ppparams$threshold_3b * 100`%**)

To alter the threshold use  parameter `filter_compound_bs_max_mv_ratio`.

Number of compounds before: **`r length(unique(datasets$filter_compounds_by_sample_all_mv_kept$Compound))`**
Number of compounds left: **`r length(unique(datasets$filter_compounds_by_sample_mv_kept$Compound))`**

```{r}
easy_datatable(
  datasets$filter_compounds_by_sample_mv_removed, show_type = "statistics",
  caption = paste0("Removed compounds due to ", ppparams$threshold_3b*100, "% Sample MVs"))
easy_datatable(
  datasets$filter_compounds_by_sample_mv_kept,
  caption = paste0("Full data kept after ", ppparams$threshold_3b*100, "% Sample MVs filtering"))
```


## Step **3c**

3. **Remove underrepresented __compounds__** (biological filters)
    c. Based on %RSD in biological samples (**<`r ppparams$threshold_3c`%**)

To alter the threshold use  parameter `filter_compound_bs_min_rsd`.

Number of compounds before: **`r length(unique(datasets$filter_compounds_by_sample_mv_kept$Compound))`**
Number of compounds left: **`r length(unique(datasets$filter_compounds_by_sample_rsd_kept$Compound))`**

```{r}
easy_datatable(
  datasets$filter_compounds_by_sample_rsd_removed, show_type = "statistics",
  caption = paste0(
    "Removed compounds due to ", ppparams$threshold_3c, " %RSD in biological samples")
  )
easy_datatable(
  datasets$filter_compounds_by_sample_rsd_kept,
  caption = paste0(
    "Full data kept after ", ppparams$threshold_3c, " %RSD filtering in biological samples")
  )
```


## Step **4**

4. **Remove underrepresented __biological samples__** (biological filters)
    a. Based on missing values ratio of compounds (**>=`r ppparams$threshold_4 * 100`%**)

To alter the threshold use parameter `filter_sample_max_mv_ratio`.

Number of all samples before: **`r length(unique(datasets$filter_compounds_by_sample_rsd_kept$Sample.Name))`**
Number of all samples left: **`r length(unique(datasets$filter_samples_by_compound_mv_kept$Sample.Name))`**

Number of biological samples before:
**`r datasets$filter_compounds_by_sample_rsd_kept %>% filter(Sample.Type == SAMPLE_TYPE_BIOLOGICAL) %>% distinct(Sample.Name) %>% nrow()`**
Number of biological samples left:
**`r datasets$filter_samples_by_compound_mv_kept %>% filter(Sample.Type == SAMPLE_TYPE_BIOLOGICAL) %>% distinct(Sample.Name) %>% nrow()`**

```{r}
easy_datatable(
  datasets$filter_samples_by_compound_mv_removed, show_type = "statistics",
  caption = paste0("Removed samples due to ", ppparams$threshold_4*100, "% Compound MVs"))
easy_datatable(
  datasets$filter_samples_by_compound_mv_kept,
  caption = paste0("Full data kept after ", ppparams$threshold_4*100, "% Compound MVs filtering"))
```


## Original

Missing value counts either by samples or by compounds based on unprocessed data (i.e. no values
have been turned into missing values based on status and nothing is filtered) as well as the full
unprocessed dataset is available in the tables below.

```{r}
easy_datatable(
  summarize_missing_values_per_sample(datasets$original, plot = FALSE),
  round_colums = "% Missing Values",
  caption = "Missing values per sample before status preprocessing",
  show_type = "statistics"
)

easy_datatable(
  summarize_missing_values_per_compound(datasets$original, plot = FALSE),
  round_colums = "% Missing Values",
  caption = "Missing values per compound before status preprocessing",
  show_type = "statistics"
)

easy_datatable(
  datasets$original,
  caption = paste0("Full data before preprocessing"))
```
