## Restructuring

<!-- Most of the restructuring is currently done already during the import of Biocrates data. -->

The imported data was restructured as follows:

* Convert to **long table** (or [molten dataset](http://dx.doi.org/10.18637/jss.v059.i10)). Each row
includes only one measurement, i.e. the data of one compound in one sample and all corresponding
information.
* Add **Sample Identification "Blank"** for blank samples, as they are empty when exported from
MetIDQ.
* Add **Sample Type "`r SAMPLE_TYPE_POOLED_QC`"** for pooled QC samples. `r SAMPLE_TYPE_POOLED_QC`
samples are identified by any occurrence of the term "pool" (case-insensitive) in the column
`r params$pool_indicator` (parameter `pool_indicator`). In contrast, **"Reference QC"** will refer
to Biocrates' QC Level 2 samples.
* Add **Sequence Position**, i.e. sample acquisition order. This position is calculated from
MetIDQ's Well Position, as this one depicts a position by
[row-major order](https://en.wikipedia.org/wiki/Row-_and_column-major_order) which does not reflect
the actual acquisition order.
* Add **Well Coordinates**, indicating the two dimensional position on the well plate. This position
is calculated from MetIDQ's Well Position and indicates rows by letters and columns by numbers.
* Add a **unique Sample Name**, which is a combination of Sample Identifier, Sequence Position and
potentially Batch (if available). These names assure unambiguous identification and analysis of
measurements within or between several batches (e.g. in particular technical samples such as
calibration standard or reference QC samples share sample identifications).
* Remove **nonessential columns**, as some of them currently interfere with proper data merging or
processing later. This includes (for now, to be optimized/reduced):
Plate Bar Code,
Sample Bar Code,
Submission Name,
Material,
Plate Production No.,
Plate Note,
Run Number,
Injection Number,
Measurement Time,
Sample Description,
Collection Date,
Org. Info and
OP.
* Unify **missing values**, i.e. "" (empty field), "0" (zero) and "NA" (not available) are converted
to "NA".
* **Factorize Compounds** with level order based on compound class and compound name.


```{r filter, results="asis"}
# Filter predefined samples
removed_samples <- data.frame()
if ("filter_regex" %in% names(params) && params$filter_regex != "") {
  # Print samples to be removed
  cat("### Sample filtering\n")
  cat(paste0("Samples removed due to regular expression filter `", params$filter_regex, "`:\n"))
  removed_samples <- biocrates %>%
    filter(grepl(params$filter_regex, Sample.Identification)) %>%
    select(Sample.Identification, Batch) %>%
    unique() %>%
    arrange(Batch, Sample.Identification)

  # Remove samples
  biocrates <- biocrates %>%
    filter(!grepl(params$filter_regex, Sample.Identification))
}
easy_datatable(removed_samples, caption = "Removed samples", show_type = "statistics")
```

```{r}
# Factorize compounds ordered by compound class and name
compound_levels <- biocrates %>%
  arrange(Class, Compound) %>%
  select(Compound) %>%
  distinct() %>%
  pull(Compound)
biocrates$Compound <- factor(biocrates$Compound, levels = compound_levels)
```

```{r, results="asis"}
# Are there any SAMPLE_TYPE_POOLED_QC samples? More than one?
POOLS_AVAILABLE <- SAMPLE_TYPE_POOLED_QC %in% biocrates$Sample.Type
POOLS_ENOUGH <- biocrates %>% filter(Sample.Type == SAMPLE_TYPE_POOLED_QC) %>%
  select(Sample.Name) %>% distinct() %>% pull(Sample.Name) %>% length() > 1
```
